---
title: "Karganov expedition analysis"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is an analysis of the beta diversity and bristol score data obtained from the Karganov expedition. 
Gut microbiota was tracked in 9 members of the Northern expedition crew. The participants included those who were already in the sea (**veterans** - 5 persons) and newly connected (**newbies** - 4 persons). Each of them had a stool sample taken before and after the expedition. So the size of all samples is 18. Further in the text **n** will denote newbies, **v** -- veterans, **0** -- before and **1** -- after. Thus n0, for example, will denote a group of samples collected from newbies before the expedition.

## Beta diversity analysis:
Beta-diversity file is numeric table 18x18 with pairwise distances between samples.
It is important that in the table first go n0 then v0 then n1 then v1.

Reading the data: 
```{r}
data <-read.csv("/home/maria/bioinf/expedition/beta/beta_diversity.csv")
dt <- as.matrix(data[1:18, 2:19])
```
Analysis of the distance in subject before and after the expedition in a group of newbies:
```{r}
newbie_ba <- c(dt[1,10], dt[2,11], dt[3,12], dt[4,13])
summary(newbie_ba)
```
Analysis of the distance in subject before and after the expedition in a group of veterans:
```{r}
veterans_ba <-c(dt[5,14], dt[6,15], dt[7,16], dt[8,17], dt[9,18])
summary(veterans_ba)
```
Wilcoxon test between these two groups:
```{r}
wilcox.test(newbie_ba, veterans_ba, paired = FALSE,  alternative = "g")
```
Levene Test computes for homogeneity of variance across these groups:
```{r}
library('car')
leveneTest(c(newbie_ba, veterans_ba),factor(c(1,1,1,1,2,2,2,2,2)))
```

Allocation of various groups from the beta-diversity table (for example, the group n0_n1 consists of pairwise distances between samples of newbies before and newbies after the expedition):
```{r}
n0_v1 <- c(dt[1:4, 14:18])
n0_v0 <- c(dt[1:4, 5:9])
n0_n1 <- c(dt[1:4, 10:13])
n0_n1_without_diagonal <- n0_n1[c(c(2:5),c(7:10),c(12:15))]
n0_n0 <- c(dt[1:4,1:4])
n0_n0 <- n0_n0[c(c(2:5),c(7:10),c(12:15))]
n1_n1 <- c(dt[10:13, 10:13])
n1_n1 <- n1_n1[c(c(2:5),c(7:10),c(12:15))]
v0_v0 <- c(dt[5:9, 5:9])
v0_v0 <- v0_v0[c(c(2:6),c(8:12),c(14:18),c(20:24))]
v1_v1 <- dt[14:18,14:18]
v1_v1 <- v1_v1[c(c(2:6),c(8:12),c(14:18),c(20:24))]
n1_v1 <- c(dt[10:13, 14:18])
n1_v0 <- c(dt[10:13, 5:9])
v0_v1 <- c(dt[5:9, 14:18])
v0_v1_without_diagonal <- v0_v1[c(c(2:6),c(8:12),c(14:18),c(20:24))]
```

Boxplot of all groups:
```{r, echo=FALSE}
library(ggplot2)
groups <- c(rep("v0_v0",20), rep("n0_n0",12),rep("n0_n1",12), rep("n0_v1", 20), rep("n0_v0", 20), rep("n1_v1", 20),
            rep("n1_v0", 20), rep("v0_v1", 20),rep("v1_v1",20), rep("n1_n1",12))
beta_divertsity <- c( v0_v0, n0_n0,n0_n1_without_diagonal, n0_v1, n0_v0, n1_v1, n1_v0, v0_v1_without_diagonal, v1_v1, n1_n1)
dt <- data.frame(gp=groups, bd=beta_divertsity, or=c(rep(0,20),rep(1,12),rep(2,12), rep(3, 20), rep(4, 20),
                                                     rep(5, 20), rep(6, 20), rep(7, 20),rep(8,20), rep(9,12)))
p <- ggplot(dt, aes(x = reorder(gp, or), y=bd)) + 
  geom_boxplot()
p + geom_dotplot(binaxis='y', stackdir='center', dotsize=0.4) + xlab("groups") + ylab("beta diversity")

```

Wilcoxon test between n0_n0 and n1_n1 (significant p-value is responsible for the convergence of newbiers after the expedition):
```{r}
wilcox.test(n0_n0, n1_n1, paired = FALSE, alternative = 'g')
```

## Bristol score analysis:

The bristol score data:
```{r}
before <- c(4,6,4,2,4,7,4,2,4)
after <- c(4,4,4,4,4,5,4,4,3)
```

F test to compare two variance:
```{r}
var.test(before,after)
```

Boxplots of Bristol score before and after the expedition (green points are newbies, blue points are veterans). 
```{r, echo = FALSE}
groups <- factor(c(rep(" before",9),rep("after",9)))
bristol_score <- c(before, after)
bristol_plot <- ggplot(data=NULL, aes(groups, bristol_score)) + 
  geom_boxplot()
color = c(3, rep(4,4), 3,3,3,4, rep(4,4), rep(3,4),4)
bristol_plot + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.5, colour = color) + ylab("bristol score")

```

Bristol score in the different groups:
```{r}
bristol_score_n0 <- c(4,6,4,2)
bristol_score_v0 <- c(4,7,4,2,4)
bristol_score_n1 <- c(4,4,4,4)
bristol_score_v1 <- c(4,5,4,4,3)
```
F test to compare variance of newbie's bristol score before and after expedition:
```{r}
var.test(bristol_score_n0, bristol_score_n1)
```
F test to compare variance of veteran's bristol score before and after expedition:
```{r}
var.test(bristol_score_v0, bristol_score_v1)
```